---
title: "Practicing Mapping"
author: "Abby Bergman"
date: "11/14/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load packages
library(tidyverse)
library(ggmap)
library(RColorBrewer)
library(gridExtra)
library(fiftystater)
library(rscorecard)
library(knitr)
library(dplyr)
library(dbplyr)
library(RSQLite)
library(sf)
library(viridis)
```


#USA data
data("fifty_states")
as_tibble(fifty_states) 

#make simple feature
st_as_sf(fifty_states, coords = c("long", "lat"))%>% 
  
  # convert fifty_states to an sf data frame
  (sf_fifty <- st_as_sf(fifty_states, coords = c("long", "lat")) %>% 
     # convert sets of points to polygons
     group_by(id, piece) %>% 
     summarize(do_union = FALSE) %>%
     st_cast("POLYGON") %>%
     # convert polygons to multipolygons for states with discontinuous regions
     group_by(id) %>%
     summarize())

ggplot(data = sf_fifty) +
  geom_sf()




```{r}

#import school data
schools_by_state <- read_csv("/Users/AbigailBergman/Desktop/Grad School/Fall Quarter 2018/Computing/Homework/hw7/schools_by_state.csv")

population <- read_csv("/Users/AbigailBergman/Desktop/Grad School/Fall Quarter 2018/Computing/Homework/hw7/population.csv")


# convert fifty_states to an sf data frame
(sf_fifty <- st_as_sf(fifty_states, coords = c("long", "lat")) %>% 
   # convert sets of points to polygons
   group_by(id, piece) %>% 
   summarize(do_union = FALSE) %>%
   st_cast("POLYGON") %>%
   # convert polygons to multipolygons for states with discontinuous regions
   group_by(id) %>%
   summarize())

#need to do left join
combined1 <-left_join(schools_by_state %>% mutate(id = tolower(id)), sf_fifty, by = "id") %>%
  filter(abb != "AK") %>%
  filter(abb != "HI")

combined <-left_join(combined1, population %>% mutate(id = tolower(id)), by = "id")

combined <- combined %>%
  mutate(per_capita = total/`2018 Population`)

#graph

ggplot(data = combined) +
  geom_sf(aes(fill = per_capita))

ggplot(data = combined1)+
  geom_sf(aes(fill = total)) + 
  labs(title = "Map of USA, Number of Colleges by State")

#bin data
combined %>%
  mutate(capita_cut = cut_number(per_capita, 4))%>%
  ggplot() +
  geom_sf(aes(fill = capita_cut)) +
  labs(title = "Number of Universities", 
       subtitle = "Divided by State Population") +
  guides(fill=guide_legend(title="Number of Universities Per Person"))

combined1 %>%
  mutate(total_cut = cut_number(total, 4))%>%
  ggplot() +
  geom_sf(aes(fill = total_cut)) +
  labs(title = "Number of Universities in each State")


display.brewer.all(type = "div")

```


```{r}
#importing college data

col_shape <- st_read("/Users/AbigailBergman/Desktop/Grad School/Fall Quarter 2018/Computing/Homework/hw7/CollegesUniversities/CollegesUniversities.shp")

#get rid of alaska and hawaii
college_shape <- col_shape %>%
  filter(COUNTRY == "US") %>%
  filter(LSTATE != "HI") %>%
  filter(LSTATE != "AK")

ggplot(data = college_shape) +
  geom_sf()

st_as_sf(fifty_states, coords = c("long", "lat")) %>% 
  # convert sets of points to polygons
  group_by(id, piece) %>% 
  summarize(do_union = FALSE) %>%
  st_cast("POLYGON")%>%
  group_by(id) %>%
  summarize()

st_crs(sf_fifty) <- 4326

ggplot(data = combined1) +
  geom_sf() +
  geom_point(data = college_shape, aes(LON, LAT)) 

#map with total by color and points for enrollment
combined1 %>%
  mutate(total_cut = cut_number(total, 4))%>%
  ggplot() +
  geom_sf(aes(fill = total_cut)) +
  labs(title = "Number of Universities in each State") + 
  geom_point(data = college_shape, aes(x = LON, y = LAT, size = TOT_ENROLL), shape = 1, fill = "grey", alpha = .15) + 
  guides(fill=guide_legend(title="Total Number of Universities", size = guide_legend(title = "Total Enrollment"))) + scale_color_brewer()

combined %>%
  mutate(capita_cut = cut_number(per_capita, 4))%>%
  ggplot() +
  geom_sf(aes(fill = capita_cut)) +
  labs(title = "Number of Universities", 
       subtitle = "Divided by State Population") +
  guides(fill=guide_legend(title="Number of Universities Per Person"))

#blank map with colleges
ggplot(data = sf_fifty) +
  geom_sf() +
  labs(title = "Number of Universities in each State") + 
  geom_point(data = college_shape, aes(x = LON, y = LAT, size = TOT_ENROLL), alpha = .1, shape = 1) 
```
For this assignment, I was interested in exploring the density of universities in different states. At first, I created several maps to see which one would provide the ideal way to visualize my data. The final map shows the contiguous United States, colored and overlayed with points. The colors show the total number of universities in a given state, split into four groups. The points show the locations of universities, with size of the point denoting total enrollment of the university. According to this map, states including California, Texas and several eastern states. This map draws an interesting contrast with the other map because the states that have the highest ratio of universities to population occures in states such as Montana and Maine. 










