---
title: "Practicing Mapping"
author: "Abby Bergman"
date: "11/14/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
#load packages
library(tidyverse)
library(ggmap)
library(RColorBrewer)
library(gridExtra)
library(fiftystater)
library(rscorecard)
library(knitr)
library(dplyr)
library(dbplyr)
library(RSQLite)
library(sf)
library(viridis)
```


```{r, include = FALSE}
#import school data
schools_by_state <- read_csv("/Users/AbigailBergman/Desktop/Grad School/Fall Quarter 2018/Computing/Homework/hw7/schools_by_state.csv")

population <- read_csv("/Users/AbigailBergman/Desktop/Grad School/Fall Quarter 2018/Computing/Homework/hw7/population.csv")


# convert fifty_states to an sf data frame
(sf_fifty <- st_as_sf(fifty_states, coords = c("long", "lat")) %>% 
   # convert sets of points to polygons
   group_by(id, piece) %>% 
   summarize(do_union = FALSE) %>%
   st_cast("POLYGON") %>%
   # convert polygons to multipolygons for states with discontinuous regions
   group_by(id) %>%
   summarize())

#need to do left join
combined1 <-left_join(schools_by_state %>% mutate(id = tolower(id)), sf_fifty, by = "id") %>%
  filter(abb != "AK") %>%
  filter(abb != "HI")

combined <-left_join(combined1, population %>% mutate(id = tolower(id)), by = "id")

combined <- combined %>%
  mutate(per_capita = total/`2018 Population`)

#graph

ggplot(data = combined) +
  geom_sf(aes(fill = per_capita))

ggplot(data = combined1)+
  geom_sf(aes(fill = total)) + 
  labs(title = "Map of USA, Number of Colleges by State")

#bin data
combined %>%
  mutate(capita_cut = cut_number(per_capita, 4))%>%
  ggplot() +
  geom_sf(aes(fill = capita_cut)) +
  labs(title = "Number of Universities", 
       subtitle = "Divided by State Population") +
  guides(fill=guide_legend(title="Number of Universities Per Person"))

combined1 %>%
  mutate(total_cut = cut_number(total, 4))%>%
  ggplot() +
  geom_sf(aes(fill = total_cut)) +
  labs(title = "Number of Universities in each State")

```


```{r, include = FALSE}
#importing college data

col_shape <- st_read("/Users/AbigailBergman/Desktop/Grad School/Fall Quarter 2018/Computing/Homework/hw7/CollegesUniversities/CollegesUniversities.shp")

#get rid of alaska and hawaii
college_shape <- col_shape %>%
  filter(COUNTRY == "US") %>%
  filter(LSTATE != "HI") %>%
  filter(LSTATE != "AK")

ggplot(data = college_shape) +
  geom_sf()
```

```{r, include=FALSE}

st_as_sf(fifty_states, coords = c("long", "lat")) %>% 
  # convert sets of points to polygons
  group_by(id, piece) %>% 
  summarize(do_union = FALSE) %>%
  st_cast("POLYGON")%>%
  group_by(id) %>%
  summarize()

st_crs(sf_fifty) <- 4326

ggplot(data = combined1) +
  geom_sf() +
  geom_point(data = college_shape, aes(LON, LAT)) 
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#map with total by color and points for enrollment
combined1 %>%
  mutate(total_cut = cut_number(total, 4))%>%
  ggplot() +
  geom_sf(aes(fill = total_cut)) +
  labs(title = "Map 1: Number of Universities in each State") + 
  geom_point(data = college_shape, aes(x = LON, y = LAT, size = TOT_ENROLL), fill = "grey", alpha = .2) + 
  guides(fill=guide_legend(title="Total Number of Universities"), size=guide_legend(title = "University Enrollment"))+ 
   scale_size(range = c(0,10)) 

combined %>%
  mutate(capita_cut = cut_number(per_capita, 4))%>%
  ggplot() +
  geom_sf(aes(fill = capita_cut)) +
  labs(title = "Map 2: Number of Universities", 
       subtitle = "Divided by State Population") +
  guides(fill=guide_legend(title="Number of Universities Per Person"))

#blank map with colleges
sf_fifty %>%
   filter(id != "alaska") %>%
  filter(id != "hawaii") %>%
ggplot() +
  geom_sf() +
  labs(title = "Map 3: Location of Universities in each State") + 
  geom_point(data = college_shape, aes(x = LON, y = LAT, size = TOT_ENROLL), alpha = .2) + 
  guides(size=guide_legend(title="University Enrollment")) +
  scale_size(range = c(0,10))
```

For this assignment, I was interested in exploring the density of universities in different states. At first, I created several maps to see which one would provide the ideal way to visualize my data. The Map 1 shows the contiguous United States, colored and overlayed with points. The colors show the total number of universities in a given state, split into four groups. The points show the locations of universities, with size of the point denoting total enrollment of the university. According to this map, states including California, Texas and several eastern states. 

This map draws an interesting contrast with the Map 2, which shows the number of universities in each state divided by state population, because the states that have the highest ratio of universities to population occurs in states such as Montana and Maine. I also chose to display map 3 which only shows the university location (points) and total enrollment (point size).

The range of point sizes was chosen so that differences in enrollment size would be clearly defined. I chose to display the number of universities per state as cut intervals because when they were initially displayed as continuous, it was difficult to distinguish between the values. Showing them in cut intervals made it easier to visualize states as having high, moderate, or low numbers of universities. 










